---
title: "Explorer R 101"
author: "Binesh Kumar"
date: "6/24/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Data Analysis Using R

This is a code walk through on some of the key features of **R** 

The session was initialy designed for a 3 hour workshop, however because of Work From Home situation condensed to walk-through session.

Please feel free to reach out to Binesh Kumar for any questions.

## What is R?

- "R is an open source software environment for statistical computing and graphics" - http://www.r-project.org
- It is a full-featured statistical analysis package (e.g., like SAS, Stata, SPSS, etc.) that allows you to:
-- Import and export data in a wide range of formats
-- Manipulate data
-- Analyse data
-- Graphically represent data
- It is a full-featured programming language designed for data analysis
- Why the name "R"?
-- First letter of two originators: Ross Ihaka and Robert Gentleman
-- Built on a earlier language called "S"


## Basic Arithmetic and Logical Operations
- You can use R like a calculator 

```{r cars, echo = TRUE}

1 + 1 # addition

10 - 9 # subtraction

10 * 10 # multiplication

100 / 10 # division

10 ^ 2 # exponentiation

abs(-10) # absolute value

ceiling(3.5) # round up to next integer

floor(3.5) # round down to next integer

sqrt(100) # square roots

exp(2) # exponential

pi # mathematical constant pi

exp(1) # mathematical constant e

log(100) # natural logs (i.e., base e)

log(100, base= 10) # base 10 logs
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

